# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r,echo=TRUE}
activity <- read.table("activity.csv",header=TRUE,sep=",")
```
## What is mean total number of steps taken per day?
### Histogram of the Total number of steps per day:
```{r,echo=TRUE}
hist(tapply(activity$steps,activity$date,sum,na.rm=TRUE),main="Total steps per day",xlab="Steps")
mdn <- median(tapply(activity$steps,activity$date,sum,na.rm=TRUE))
mn <- mean(tapply(activity$steps,activity$date,sum,na.rm=TRUE))
```
### Mean per day: `r mn`
### Median per day: `r mdn`

## What is the average daily activity pattern?
### Plot of the  5-minute interval / Average number of steps(across all days):
```{r,echo=TRUE}
ave_steps <- aggregate(steps ~ interval, data=activity, mean)
plot(ave_steps$interval,ave_steps$steps,type="l",xlab="5-minute intervals",ylab="Average steps",main="Average steps per 5-minute intervals")

max_interval<-ave_steps$interval[which.max(ave_steps$steps)]
average_steps_max<-ave_steps$steps[which.max(ave_steps$steps)]
```
### 5-minute interval with maximun number of steps on average: `r max_interval`
### Maximun number of steps on average: `r average_steps_max`
## Imputing missing values
```{r,echo=TRUE}
miss<-sum(is.na(activity$steps))
```
### Total number of missing values in datasets is : `r miss`
### The strategy for filling in all of the missing values in the dataset consist in to use the mean for that 5-minute interval.
```{r,echo=TRUE}
activity <- merge(activity, ave_steps, by="interval", suffixes=c("","_meanxinterval"))
activity[is.na(activity$steps),2]<-activity[is.na(activity$steps),4]

hist(tapply(activity$steps,activity$date,sum,na.rm=TRUE),main="Total steps per day (missing values off)",xlab="Steps")
mdn <- median(tapply(activity$steps,activity$date,sum,na.rm=TRUE))
mn <- mean(tapply(activity$steps,activity$date,sum,na.rm=TRUE))
```
### Mean per day: `r mn`
### Median per day: `r mdn`

### These values differ from the estimates from the first part of the assignment, the mean and median are the same. Additionally the histogram is more symmetric around de mean/median. The impact of imputing missing data on the estimates of the total daily number of steps is high.


## Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}
# Set language to English
Sys.setenv(LANGUAGE="en")
Sys.setlocale("LC_TIME", "English")

# Create column with day of week
activity$dayOfWeek<-as.factor(weekdays(as.Date(activity$date)))

# Create column with type day of week
activity$typeDayOfWeek <- as.factor(sapply(activity$dayOfWeek,function(day) if (day %in% c("Saturday","Sunday")){ "weekend" }else {"weekday"}))

# Average steps for weekend
ave_steps_weekend <- aggregate(steps ~ interval,data=activity,subset=activity$typeDayOfWeek=="weekend",mean)

# Average steps for weekday
ave_steps_weekday <- aggregate(steps ~ interval,data=activity,subset=activity$typeDayOfWeek=="weekday",mean)

# Plots Average steps for weekend and Average steps for weekday
par(mfrow=c(2,1))

plot(ave_steps_weekend$interval,ave_steps_weekend$steps,type="l",xlab="5-minute intervals",ylab="Average steps",main="Average steps per 5-minute intervals in weekend")
plot(ave_steps_weekday$interval,ave_steps_weekday$steps,type="l",xlab="5-minute intervals",ylab="Average steps",main="Average steps per 5-minute intervals in weekday")

```
### There are more average steps in weekend compared to weekdays.